---
title: "Analise De Agrupamentos"
author: "gsh87"
date: "maio 2018"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
    smooth_scroll: TRUE
---

<style type="text/css">

.chart-title {  /* chart_title  */
   font-size: 20px;
</style>


<!-- ################################################# -->

```{r,echo=FALSE,include=FALSE}
################################### 0 - Limpeza
# removendo lixo da memoria
rm(list=ls())
set.seed(123)
```

```{r,echo=FALSE,results='hide',warning=FALSE,message=FALSE}
## 1 - Pacotes
options(repos=c(CRAN="http://vps.fmvz.usp.br/CRAN/"))
library(pacman)    # pacote para load das bibliotecas
p_load(stats)     #  Modelos ARIMA
p_load(agricolae)# testes
p_load(lmtest)    # testes de residuos
p_load(randtests) # pacotes de testes estatisticos
p_load(rms)       # pacote para stepwise
p_load(forecast)  # forecast
p_load(smooth)    # SMA - simple moving average
p_load(nnfor)     # neural networks for time series
p_load(h2o)
p_load(fBasics)   # estatistica basica

p_load(fpc)      # Flexible Procedures for Clustering
p_load(factoextra) #factoextra for visualizing clusters using ggplot2 plotting system
p_load(cluster)   # cluster for computing pam and for analyzing cluster silhouettes
p_load(NbClust) #NbClust for finding the optimal number of clusters
p_load(factoextra) # numero otimo de clusters

## pacotes graficos ##
p_load(ggplot2)   # grafico
p_load(ggthemes)  # grafico 
p_load(gridExtra) # Graficos em matrizes
p_load(plotly)    # interative graphics
p_load(ggcorrplot) # correlacao
p_load(ggExtra)     # Add marginal Histogram
p_load(ggalt)        # circulos no ggplot

## pacotes de manipulacao ##
p_load(lubridate) # tratamento de datas
p_load(plyr)     # empilhamento
p_load(magrittr)   # pipeline
p_load(dplyr)      # manipulacao de base
p_load(knitr)       # apresentacao
p_load(binomTools)   # extrai
p_load(vcd)          
p_load(DescTools)      # summary grafico
p_load(RODBC)          # conexao ODC
p_load(sqldf)          # sqldf
p_load(corrplot)       # correlacao
p_load(prais)
p_load(Metrics)
p_load(dummies)
p_load(reshape2)       # empilha as colunas usando uma flag

## Packages visual
p_load(flexdashboard)
p_load(DT)


p_load(factoextra) #factoextra for visualizing clusters using ggplot2 plotting system
p_load(cluster)   # cluster for computing pam and for analyzing cluster silhouettes
p_load(NbClust) #NbClust for finding the optimal number of clusters

```

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
p_load(datasets)

Base_iris <- iris %>% mutate(
  flag_Sep_comprimento = ifelse(Sepal.Length>=(mean(Sepal.Length)+sd(Sepal.Length)),1,0),
  flag_Sep_largura = ifelse(Sepal.Width>=(mean(Sepal.Width)+sd(Sepal.Width)),1,0)) %>% `colnames<-`(c("Sepal_Length","Sepal_Width","Petala_Length","Petala_Width","Classificacao","flag_Sep_comprimento","flag_Sep_largura"))



```



# Agrupamentos Kmeans {data-orientation=rows}

Row 
-------------------------------------
### Cluster: Visualizacao

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
df <- iris[c(1, 2, 3, 4)]
pca_mod <- prcomp(df)  # compute principal components

# Data frame of principal components ----------------------
df_pc <- data.frame(pca_mod$x, Species=iris$Species)  # dataframe of principal components
df_pc_vir <- df_pc[df_pc$Species == "virginica", ]  # df for 'virginica'
df_pc_set <- df_pc[df_pc$Species == "setosa", ]  # df for 'setosa'
df_pc_ver <- df_pc[df_pc$Species == "versicolor", ]  # df for 'versicolor'
 
# Plot ----------------------------------------------------
Grafico15 <- ggplot(df_pc, aes(PC1, PC2, col=Species)) + 
  geom_point(aes(shape=Species), size=2) +   # draw points
  labs(title="Clusterização", 
       subtitle="Clusterização com componentes principais",
       caption="Source: Iris") + 
  coord_cartesian(xlim = 1.2 * c(min(df_pc$PC1), max(df_pc$PC1)), 
                  ylim = 1.2 * c(min(df_pc$PC2), max(df_pc$PC2))) +   # change axis limits
  geom_encircle(data = df_pc_vir, aes(x=PC1, y=PC2)) +   # draw circles
  geom_encircle(data = df_pc_set, aes(x=PC1, y=PC2)) + 
  geom_encircle(data = df_pc_ver, aes(x=PC1, y=PC2))

Grafico15
```

Row {.tabset .tabset-fade}
-------------------------------------
### Cluster: Kmeans
```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}


Base_Cluster <- Base_iris %>% select("Sepal_Length","Sepal_Width","Petala_Length","Petala_Width") %>% scale()
# K-Means Clustering with 5 clusters
fit <- kmeans(Base_Cluster, 3)

# Cluster Plot against 1st 2 principal components

# vary parameters for most readable graph
require(cluster) 
clusplot(Base_Cluster, fit$cluster, color=TRUE, shade=TRUE, 
  	labels=2, lines=1,main="Clusterizacao")

```

### Kmeans Impressão dos Grupos

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
# Centroid Plot against 1st 2 discriminant functions
require(fpc)
plotcluster(Base_Cluster, fit$cluster,main="Grupos")
```

###  Analise do numero otimo de clusters

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
set.seed(123)
#total within-cluster sum of square (wss) 
# Compute and plot wss for k = 2 to k = 15
k.max <- 10 # Maximal number of clusters
data <- Base_Cluster
wss <- sapply(1:k.max,function(k){kmeans(data, k, nstart=10 )$tot.withinss})

plot(1:k.max, wss,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")
abline(v = 3, lty =2)
```


###  Numero otimo de clusters: factoextra wss

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
set.seed(123)
require(factoextra)
grafico_wss <- fviz_nbclust(Base_Cluster, kmeans, method = "wss") +
    geom_vline(xintercept = 3, linetype = 2)
grafico_wss
```


###  Numero otimo de clusters: factoextra Silhoeta

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
set.seed(123)
require(factoextra)
grafico_wss2 <- fviz_nbclust(Base_Cluster, kmeans, method = "silhouette") 
grafico_wss2
```

###  Numero otimo de clusters: nbClust

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE,include = FALSE}
set.seed(123)
require(factoextra)
require(NbClust)
nb <- NbClust(Base_Cluster, distance = "euclidean", min.nc = 2,
        max.nc = 10, method = "complete", index ="all")
```


```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
set.seed(123)
require(factoextra)
require(NbClust)

Grafico_nb <- fviz_nbclust(nb) + theme_minimal()

```

###  Numero otimo de clusters: nbClust Grafico

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
set.seed(123)
require(factoextra)
require(NbClust)

Grafico_nb

```

Row {.tabset .tabset-fade}
-------------------------------------
### Agrupamentos Hierarquicos

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
Base_Hierarquico <- Base_iris %>% select("Sepal_Length","Sepal_Width","Petala_Length","Petala_Width") %>% scale()

# Compute pairewise distance matrices
dist.res <- dist(Base_Hierarquico, method = "euclidean")
# Hierarchical clustering results
hc <- hclust(dist.res, method = "complete")
# Visualization of hclust
plot(hc, labels = FALSE, hang = -1)
# Add rectangle around 3 groups
rect.hclust(hc, k = 3, border = 2:4) 


#Cut into 3 groups
rotulos <- cutree(hc, k = 3)

```

###  Numero otimo de clusters: factoextra

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
set.seed(123)
require(factoextra)
grafico_hcut <- fviz_nbclust(Base_Hierarquico, hcut, method = "wss") +
    geom_vline(xintercept = 3, linetype = 2)
grafico_wss
```


###  Numero otimo de clusters: factoextra Silhoeta

```{r,echo=FALSE,warning=FALSE,fig.align='center',warning=FALSE,message=FALSE}
set.seed(123)
require(factoextra)
grafico_wss3 <- fviz_nbclust(Base_Hierarquico, kmeans, method = "silhouette") 
grafico_wss3
```



# Referência {data-orientation=rows}

Row 
-------------------------------------


* [Graficos](http://http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)

* [Clusters](http://www.sthda.com/english/wiki/print.php?id=239)
  

